---
title: "Homework 5 -- Intro to Probability and Statistics"
author: "Leticia Rodrigues"
output: pdf_document
---

# Voting in the UN General Assembly
## Question 1

We begin by examining how the distribution of state ideal points has changed since the end of communism. Plot the distribution of ideal points separately for 1980 and 2000—about 10 years before and 10 years after the fall of the Berlin Wall, respectively. Add the median to each plot as a vertical line. How do the two distributions differ? Pay attention to the degree of polarization and give a brief substantive interpretation of the results. Use the `quantile()` function to quantify the patterns you identified.

```{r}
library(ggplot2)
library(tidyverse)

unvoting <- read.csv ("https://raw.githubusercontent.com/umbertomig/intro-prob-stat-FGV/master/datasets/unvoting.csv")
head(unvoting)

plot1980 <- subset(unvoting, Year == 1980)
hist(plot1980$idealpoint,main = "Ideal Points of 1980",xlab = "Ideal Points", col = "pink")
abline(v = median(plot1980$idealpoint),col = "blue")
legend(x ="topright","Median",col="blue",lwd =1)

plot2000 <- subset(unvoting, Year == 2000)
hist(plot2000$idealpoint, main = "Ideal Points of 2000", xlab = "Ideal Points", col = "pink")
abline(v = median(plot2000$idealpoint), col = "blue")
legend(x ="topright","Median",col='blue',lwd =1)

quantile(plot1980$idealpoint)
quantile(plot2000$idealpoint)
```

Answer: The graph shows that countries constituents of UN have become less liberal over time. 

## Question 2

Next, examine how the number of countries voting with the United States has changed over time. Plot the average percentage agreement with the United States across all countries over time. Also, add the average percentage agreement with Russia as another line for comparison. Using the `tapply()` function may help with this analysis. Does the United States appear to be getting more or less isolated over time, as compared to Russia? Identify some countries that are consistently pro-US. What are the most pro-Russian countries? Give a brief substantive interpretation of the results.

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

unvoting %>% 
  group_by(Year) %>% 
  summarise(Average_US = mean(PctAgreeUS), Average_RUS = mean(PctAgreeRUSSIA)) %>%
  na.omit() %>%
  pivot_longer(cols = c(Average_US, Average_RUS)) %>%
  ggplot(aes(x = Year, y = value, colour = name)) + 
  geom_line() + 
  labs( title = "Average Percentage Agreement of US and Russia Across Countries Over Time ")

sort(tapply(unvoting$PctAgreeRUSSIA, unvoting$CountryName, mean),decreasing = TRUE) [2:8]
sort(tapply(unvoting$PctAgreeUS, unvoting$CountryName, mean), decreasing = TRUE) [2:8]
```

Answer:  According to the plot above, the idealpoint of the United States has dropped over the years, thus , the United States tends to get more isolated over time in comparison to Russia, that had an increase on the percentage of countries who has matching votes on the same issue over time.
 The countries that are most pro-Russian are te following: German Democratic Republic, Czechoslovakia, Belarus, Ukraine and PolanD, Mongolia and Hungary.
 The countries that are consistently pro-US are: Palau, United Kingdom, Taiwan, Israel, Federated States of Micronesia, Canada and Luxemburg. 
        
## Question 3

One problem with using the proportion of votes that agree with the United States or Russia as a measure of state preferences is that the ideological positions, and consequently the voting patterns, of the two countries might themselves have changed over time. This makes it difficult to know which countries’ ideological positions have changed. Investigate this issue by plotting the evolution of the two countries' ideal points over time. Add the yearly median ideal point of all countries. How might the results of this analysis modify (or not) your interpretation of the previous analysis?

```{r}
str(unvoting)

data <- filter(unvoting, CountryAbb %in% c('RUS','USA'))

Median <- group_by(unvoting, Year) %>%
  summarise(m=median(idealpoint))

ggplot(data, aes(x=Year, y=idealpoint, col=CountryName)) +
  geom_line() +
  geom_line(data=Median, aes(y=m), col="blue", lty=2) +
  labs(caption = "Comparison of idealogical positions for USA and Russia.
                  \nBlue dotted line indicates yearly median of all countries") +
  theme_minimal() +
  theme(legend.title=element_blank(), legend.position = "bottom", 
        plot.caption = element_text(hjust = 0, face="italic"))
```

Answer: The past few years indicates that the United States has remained with more liberal policies than other countries of the UN. According to the graph above, with the fall of the Soviet Union in 1991, Russia started to adopt liberal policies, thus, the country became more liberal than median UN country members. The results of this analysis modify the interpretation that I gave on question 2, because it does not mean that countries have identified with Russian policies over the years; it means that Russia has become more opened to liberal policies. 

## Question 4

Let’s examine how countries that were formerly part of the Soviet Union differ in terms of their ideology and UN voting compared to countries that were not part of the Soviet Union. The former Soviet Union countries are Estonia, Latvia, Lithuania, Belarus, Moldova, Ukraine, Armenia, Azerbaijan, Georgia, Kazakhstan, Kyrgyzstan, Tajikistan, Turkmenistan, Uzbekistan, and Russia. The `%in%` operator, which is used as `x %in% y`, may be useful. This operator returns a logical vector whose elements are `TRUE` if the corresponding element of vector x is equal to a value contained in vector y and otherwise `FALSE`. Focus on the most recently available UN data from 2012 and plot each post-Soviet Union state’s ideal point against the proportion of its votes that agree with the United States. Compare the post-Soviet Union states, within the same plot, against the other countries. Briefly comment on what you observe.

```{r}
data<- filter(unvoting, Year==2012)
svtun<- c("Estonia", "Latvia", "Lithuania", "Belarus", "Moldova", "Ukraine", "Armenia", "Azerbaijan", "Georgia", "Kazakhstan", "Kyrgyzstan", "Tajikistan", "Turkmenistan", "Uzbekistan", "Russia")

un_svt<- filter(data, CountryName %in% svtun)
un_vt_ot<- filter(data, !(CountryName %in% svtun))
  
ggplot(NULL, aes(x = PctAgreeUS, y= idealpoint, group= CountryName, color= CountryName))+
  geom_point(data= un_svt)+
    geom_point(data= un_vt_ot, alpha=1/15, color="black")+
      ggtitle("Agreement with Russia in 2012")+
        labs(y="Ideal Points", x="Agreement rate of the UN Votes", legend= "color")
```

Answer: The graph above shows that the countries that were part of the Soviet Union are close and concentrated in the middle of other countries that were not part of the country. Countries that were not part of the Soviet Union are much more spread out and dispersed. 

## Question 5

We have just seen that while some post-Soviet countries have retained nonliberal ideologies, other post-Soviet countries were much more liberal in 2012. Let’s examine how the median ideal points of Soviet/post-Soviet countries and all other countries have varied over all the years in the data. Plot these median ideal points by year. Be sure to indicate 1989, the year of the fall of the Berlin Wall, on the graph. Briefly comment on what you observe.

```{r}
Soviet_countries <- c("Estonia", "Latvia", "Lithuania", "Belarus", "Moldova", "Ukraine", "Armenia", 
                      "Azerbaijan", "Georgia", "Kazakhstan", "Kyrgyzstan", "Tajikistan", 
                      "Turkmenistan", "Uzbekistan", "Russia") 


unvoting <- mutate(unvoting, 
      svtunion=ifelse(CountryName %in% Soviet_countries, "Soviet", "non-Soviet"))

y <- aggregate(idealpoint~Year+svtunion, FUN=median, data=unvoting)

ggplot(y) +
  geom_line(aes(x=Year, y=idealpoint, col=svtunion)) +
  scale_color_discrete(name="Legend") +
  labs(title="Variation of Median Ideal Points") +
  geom_vline(xintercept = 1989, linetype="dotted")
```

Answer: According to the graph, the fall of the Berlin wall represents a significant increase in the adoption of liberal policies by the constituent countries of the USSR. The increase was so brutal that the countries of the Soviet Union managed to overtake non-USSR countries after 1989. 

## Question 6

Following the end of communism, countries that were formerly part of the Soviet Union have become much more ideologically diverse. Is this also true of the world as a whole? In other words, do countries still divide into two ideological factions? Let’s assess this question by applying the k-means clustering algorithm to ideal points and the proportion of votes agreeing with the United States. Initiate the algorithm with just two centroids and visualize the results separately for 1989 and 2012. Briefly comment on the results.

```{r}
unvoting89 <- subset(unvoting, Year == 1989)
clst_alg <- kmeans(unvoting89[, c("idealpoint", "PctAgreeUS")], centers = 2)
unvoting89$cluster1 <- clst_alg$cluster

centroids = data.frame(clst_alg$centers)

ggplot(data = unvoting89,aes(x = idealpoint, y = PctAgreeUS)) + 
        geom_point(aes(col = as.factor(cluster1))) +
        labs(title = "Ideal Points Analysis of 1989") +
         geom_point(data = centroids,color = "black", size = 4, pch = "X")

unvoting12 <- subset(unvoting, Year == 2012)
clst_alg2 <- kmeans(unvoting12[, c("idealpoint", "PctAgreeUS")], centers = 2)
unvoting12$cluster2 <- clst_alg2$cluster

ggplot(data = unvoting12,aes(x = idealpoint, y = PctAgreeUS)) + 
        geom_point(aes(col = as.factor(cluster2))) +
        labs(title = "Ideal Points Analysis of 2012") +
        geom_point(data = centroids,color = "black", size = 4, pch = "X")
```

Answer: The chart above reveals that there was no sudden change between 1989 and 2012 between countries that agree or disagree with the United States. Thus, it is concluded that the international system remained bipolarized even with the end of the Soviet Union. 

# Success of Leader Assassination as a Natural Experiment

## Question 7

How many assassination attempts are recorded in the data? How many countries experience at least one leader assassination attempt? (The `unique()` function, which returns a set of unique values from the input vector, may be useful here.) What is the average number of such attempts (per year) among these countries?

```{r}
leaders <- read.csv("https://raw.githubusercontent.com/umbertomig/intro-prob-stat-FGV/master/datasets/leaders.csv")
head(leaders)
nrow(leaders)
length(unique(leaders$country))
mean(tapply(leaders$country, leaders$year, length))
```

Answer: There were 250 assassination attempts recorded in the data. 88 countries have experienced at least one leader assassination attempt. The average number of such attempts per year were of 2.45. 

## Question 8

Create a new binary variable named `success` that is equal to 1 if a leader dies from the attack and 0 if the leader survives. Store this new variable as part of the original data frame. What is the overall success rate of leader assassination? Does the result speak to the validity of the assumption that the success of assassination attempts is randomly determined?

```{r}
leaders$success = 0
leaders$success [leaders$result =="dies within a day after the attack"] <- 1
leaders$success [leaders$result == "dies between a day and a week"] <- 1
leaders$success [leaders$result == "dies, timing unknown"] <- 1
leaders$success [leaders$result == "dies between a week and a month"] <- 1
mean(leaders$success)
```

Answer: The overall success rate of leader assassination was of 21.6%. 

## Question 9
Investigate whether the average polity score over three years prior to an assassination attempt differs on average between successful and failed attempts. Also, examine whether there is any difference in the age of targeted leaders between successful and failed attempts. Briefly interpret the results in light of the validity of the aforementioned assumption.

```{r}
leaders9 <- select(leaders, politybefore, success)
head(leaders9)

leadersuccess <- filter(leaders9, success == 1)
head(leadersuccess)

leadersfail <- filter(leaders9, success == 0)
head(leadersfail)

mean(leadersuccess$politybefore)
mean(leadersfail$politybefore)

ages <- select(leaders, age, success)

agesfail <- filter(ages, success == 0)
head(agesfail)
mean(agesfail$age)

agessuccess <- filter(ages, success == 1)
head(agessuccess)
mean(agessuccess$age)
```

Answer: There were no significant differences between the age of the victims of the attacks.
Average age of victims from successfully attempts: 56.462
Average age of victims from failed attempts: 52.714

## Question 10

Repeat the same analysis as in the previous question, but this time using the country’s experience of civil and international war. Create a new binary variable in the data frame called `warbefore`. Code the variable such that it is equal to 1 if a country is in either civil or international war during the three years prior to an assassination attempt. Provide a brief interpretation of the result.

```{r}
leaders$warbefore <- as.numeric(leaders$interwarbefore == 1|leaders$civilwarbefore == 1)

successfull <- select(leaders, success, warbefore)
successfull <- filter(successfull, success == 1)
head(successfull)
mean(successfull$warbefore)

fail <- select(leaders, success, warbefore)
fail2 <- filter(fail, success == 0)
head(fail2)
mean(fail2$warbefore)
```

Answer: There were no significant differences between the country’s experience of civil and international war with successful murder attempts and failed murder attempts. 
Average number of successfull attempts:0.351
Average number of failed attempts: 0.372

